### 2020-07-10 abstract/IQFA 11_19/abstract_CPM.tex
tout d'un coup

\documentclass[aps,twocolumn,showpacs]{revtex4-1}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{theorem}
\usepackage{amsmath}
\usepackage{times}
\usepackage{color}
\usepackage[french]{babel}
\usepackage[T1]{fontenc}
\usepackage{ifpdf}
\ifpdf
\usepackage{epstopdf}   
\usepackage{url}
\fi

\begin{document}
\title{Sub-GHz linewidths ensembles of SiV centers in a diamond nano-pyramid revealed by charge state conversion}

\author{Cl{\'e}ment Pellet-Mary$^1$, Louis Nicolas$^1$, Tom Delord $^1$, Paul Huillery $^1$, and Gabriel H{\'e}tet$^{1}$}
\email{clement.pellet-mary@phys.ens.fr}
\affiliation{$^1$ Laboratoire de Physique de l'Ecole normale sup{\'e}rieure, ENS, Universit{\'e} PSL, CNRS,Sorbonne Universit{\'e}, Universit{\'e} Paris-Diderot, Sorbonne Paris Cit{\'e}, Paris, France
}

\begin{abstract}
\normalsize
\noindent
Efficient interfaces between two-level emitters and photons are key components of quantum networks and sensing. %Diamond negatively charged silicon vacancy centers (SiV$^-$) can be used for such applications as they are photostable and reveal a large emission in the zero phonon line. Nanostructures are required in order to avoid reflections to ensure a strong coupling between the emitters and the light field. 
SiV$^-$ centers embedded into a diamond nanostructure are promising for such purpose. We have studied photoluminescence properties at cryogenic temperatures of diamond AFM probes [1,2] and observed photochromism of a very dense ensemble of SiV$^-$ centers consisting in trapping into a dark state under resonant excitation when nitrogen impurities are present. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states.  This effect is used to perform persistent hole burning which reveals very low homogeneous broading, only twice the lifetime limit. It is promising  for quantum optics experiments and paves the way to sub-wavelength microscopy technics. %such as ground state depletion microscopy

@@

\includegraphics[width=.7\columnwidth]{fig_hyp}\\
\noindent\textbf{Figure 1.} a) Averaged PLE measured at 6K by scanning the resonant laser at 737 nm around an electronic transition transition in the presence of green laser light between scans. b) PLE spectra after three different exposure times at 6K : in black without hole burning, in blue t$_{exp}$ = 300 ms and in orange, t$_{exp}$ = 1500 ms. The dots represent raw data and the continuous lines are fitted data. Here P$_{737}$= 0.877 $\mu$W. The inset shows the width of the holes $\Gamma_h$ inferred from the fit and normalized to the homogeneous linewidth $\Gamma_e$(W= $\Gamma_h/\Gamma_e$).

\end{abstract}

\maketitle

\begin{thebibliography}{}

\bibitem{Ref1} AR. Nelz et al., Applied Physics Letters, 109, 19, 193105, (2016).

\bibitem{Ref2} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).

\end{thebibliography}
\end{document}

>>>>

\documentclass[a4paper,11pt]{article}
\linespread{1.1}
\usepackage[top=3cm, bottom=2.54cm, left=2.54cm, right=2.54cm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\begin{document}
\pagenumbering{gobble}

\Large
 \begin{center}
Sub-GHz linewidths ensembles of SiV centers in a diamond nano-pyramid revealed by charge state conversion

\hspace{10pt}

\normalsize
% Author names and affiliations
Cl{\'e}ment Pellet-Mary $^1$, Louis Nicolas $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\

\hspace{10pt}

\small  
$^1$ Laboratoire de Physique de l'Ecole normale sup{\'e}rieure, ENS, Université PSL, CNRS,Sorbonne Universit{\'e}, Universit{\'e} Paris-Diderot, Sorbonne Paris Cit{\'e}, Paris, France.\\

\end{center}

\hspace{10pt}

\normalsize

\noindent
Efficient interfaces between two-level emitters and photons are key components of quantum networks and sensing. %Diamond negatively charged silicon vacancy centers (SiV$^-$) can be used for such applications as they are photostable and reveal a large emission in the zero phonon line. Nanostructures are required in order to avoid reflections to ensure a strong coupling between the emitters and the light field. 
SiV$^-$ centers embedded into a diamond nanostructure are promising for such purpose. We have studied photoluminescence properties at cryogenic temperatures of diamond AFM probes [1,2] and observed photochromism of a very dense ensemble of SiV$^-$ centers consisting in trapping into a dark state under resonant excitation when nitrogen impurities are present. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states.  This effect is used to perform persistent hole burning which reveals very low homogeneous broading, only twice the lifetime limit. It is promising  for quantum optics experiments and paves the way to sub-wavelength microscopy technics. %such as ground state depletion microscopy

@@

%In our search for a suitable sample, we have studied in detail and at cryogenic temperature sharp single crystal diamond AFM probes from Artech Carbon. Due to the growth conditions, a lot of SiV centers are present at the apex of such a tip whose radius of curvature is as low as 10 nm [3]. They exhibit remarkable optical properties at low temperature.
%We have performed measurements with a scanning Fabry-P{\'e}rot interferometer [4]. This way, we have shown that the inhomogeneous linewidth of the ensemble of emitters is 15 GHz at 6 K. That means that the structure has a very low strain which is probably due to the slow and controlled CVD growth along the $\langle001\rangle$ axis [5].

%Moreover, when nitrogen impurities are also present, SiV centers are trapped into a dark state under resonant excitation. Its lifetime is very long (more than hours). Illumination with a green laser is needed to recover the photoluminescence. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states [6]. 
%The charge transfer dynamics enables persistent spectral hole burning which unveils the homogeneous width as shown in Figure 1. Thanks to a rate model, we estimate the homogeneous linewidth $\Gamma_e$ to be 390 MHz at 6 K [7].

%Figure
\hspace{10pt}
\begin{figure}[h!]\centering
\includegraphics[width=.7\columnwidth]{fig_hyp}\\
\noindent\textbf{Figure 1.} a) Averaged PLE measured at 6K by scanning the resonant laser at 737 nm around an electronic transition transition in the presence of green laser light between scans. b) PLE spectra after three different exposure times at 6K : in black without hole burning, in blue t$_{exp}$ = 300 ms and in orange, t$_{exp}$ = 1500 ms. The dots represent raw data and the continuous lines are fitted data. Here P$_{737}$= 0.877 $\mu$W. The inset shows the width of the holes $\Gamma_h$ inferred from the fit and normalized to the homogeneous linewidth $\Gamma_e$(W= $\Gamma_h/\Gamma_e$).
\end{figure}

%This effect reveals that the linewidth is close to be lifetime limited as required for quantum optics experiments and paves the way to sub-wavelength microscopy technics such as ground state depletion microscopy [8]. 
%
%We will also present our latest results on the modification of lifetime of solid state emitters in a half-cavity set-up.

%References
\hspace{10pt}

\noindent{[}1{]} R. Nelz et al., Applied Physics Letters, 109, 19, 193105, (2016).\\
\noindent{[}2{]} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).\\
\noindent{[}2{]} L. Nicolas et al., AIP Advances, 8, 6, 065102, (2018).\\
\end{document}

### 2020-07-10 abstract/IQFA 11_19/abstract_LN.tex
tout d'un coup

\normalsize
% Author names and affiliations
Cl{\'e}ment Pellet-Mary $^1$, Louis Nicolas $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\

@@

\noindent
Efficient interfaces between two-level emitters and photons are key components of quantum networks and sensing. %Diamond negatively charged silicon vacancy centers (SiV$^-$) can be used for such applications as they are photostable and reveal a large emission in the zero phonon line. Nanostructures are required in order to avoid reflections to ensure a strong coupling between the emitters and the light field. 
SiV$^-$ centers embedded into a diamond nanostructure are promising for such purpose. We have studied photoluminescence properties at cryogenic temperatures of diamond AFM probes [1,2] and observed photochromism of a very dense ensemble of SiV$^-$ centers consisting in trapping into a dark state under resonant excitation when nitrogen impurities are present. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states.  This effect is used to perform persistent hole burning which reveals very low homogeneous broading, only twice the lifetime limit. It is promising  for quantum optics experiments and paves the way to sub-wavelength microscopy technics. %such as ground state depletion microscopy

We will also present our latest results on the Zeeman splitting and possible lifetime measurement of the SiV$^-$ $1/2$ spin.

>>>>

\normalsize
% Author names and affiliations
Louis Nicolas $^1$, Cl{\'e}ment Pellet-Mary $^1$, Tom Delord $^1$, Paul Huillery $^1$, Gabriel H{\'e}tet$^1$\\

@@

\noindent
Efficient interfaces between two-level emitters and photons are key components of quantum networks and sensing. %Diamond negatively charged silicon vacancy centers (SiV$^-$) can be used for such applications as they are photostable and reveal a large emission in the zero phonon line. Nanostructures are required in order to avoid reflections to ensure a strong coupling between the emitters and the light field. 
SiV$^-$ centers embedded into a diamond nanostructure are promising for such purpose. We have studied photoluminescence properties at cryogenic temperatures of diamond AFM probes [1,2] and observed photochromism of a very dense ensemble of SiV$^-$ centers consisting in trapping into a dark state under resonant excitation when nitrogen impurities are present. We suppose that this effect is a charge state switching between negatively charged and neutral SiV charge states.  This effect is used to perform persistent hole burning which reveals very low homogeneous broading, only twice the lifetime limit. It is promising  for quantum optics experiments and paves the way to sub-wavelength microscopy technics. %such as ground state depletion microscopy
We will also present our latest results on the modification of the lifetime of solid state emitters in a half-cavity set-up.

### 2020-07-10 ensignement/TD_math/CC1_projet.tex
tout d'un coup

\documentclass[fleqn]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{mathexam}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{mathtools, physics}

@@

\begin{document}
\ExamInstrBox{
\begin{center}
\textbf{\textbf{Corrigé}}
\end{center}
}

\section{Exercice 1 : Intégrales }

\begin{enumerate}
   \item \textbf{(3 pts)} En intégrant par partie,
   \begin{align*}
   \int_0^1 x^2 \mathrm{ln}(x) dx = \left[\dfrac{x^3}{3} \mathrm{ln}(x)\right]_0^1 - \int_0^1 \dfrac{x^2}{3}dx   =-\left[\dfrac{x^3}{9}\right]_0^1 =-\dfrac{1}{9}
   \end{align*}
   L'intégrale n'est pas généralisée car $\lim_{x \to 0} \left(x^2 \mathrm{ln}(x)\right)= 0$ par croissance comparée, la fonction est bien bornée sur son domaine d'intégration.
   \item \textbf{(3 pts)} En posant le changement de variable $y^2=1+3\sin(x)$, qui est bien bijectif sur l'intervalle considéré, alors $2y\mathrm{d}y=3\cos(x)dx$ et les nouvelles bornes vont de $y=1$ à $y=2$.
   \begin{align*}
   \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx = \int_1^2 \dfrac{\dfrac{2}{3}y\mathrm{d}y}{\sqrt{y^2}}=\int_1^2 \dfrac{2}{3}dy=\dfrac{2}{3}
   \end{align*}
   L'intégrale n'est pas généralisée.
   \item \textbf{(4 pts)} En remarquant que $$\dfrac{\mathrm{d}(1/\sin(x))}{dx}=-\dfrac{\cos(x)}{\sin^2(x)}$$
   Et que $$ \dfrac{1}{\sin^2(x)}=\dfrac{\sin^2(x)+\cos^2(x)}{\sin^2(x)}=1+\mathrm{cotan}^2(x)=-\mathrm{cotan}'(x)$$
   Il vient que $$\int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx = \left[-\dfrac{1}{\sin(x)}+\mathrm{cotan}(x)\right]_{-\pi/2}^{\pi/2} =\left[\dfrac{\cos(x)-1}{\sin(x)}\right]_{-\pi/2}^{\pi/2}=-2$$
   L'intégrale n'est pas généralisée car $\lim_{x \to 0}\left(\dfrac{\cos(x)-1}{\sin^2(x)}\right)=-\dfrac{1}{2}$ par les développement limités de $\sin^2(x)$ et $\cos(x) -1$ en $0$. La fonction est bien bornée sur son domaine d'intégration.
\end{enumerate}
\section{Exercice 2 : calcul de volumes}
\begin{enumerate}
\item \textbf{(3 pts)} On décompose notre cylindre en petit cylindre de hauteur $\mathrm{d} z$ et additionne les volumes de tous ces cylindres pour trouver le volume du cylindre. Le volume de chacun de ces cylindres est $\pi r(z)^2 \mathrm{d} z$. Il ne reste plus qu'à trouver $r(z)$.\newline
On connaît les équations du cylindre $x^2 + y^2 = R^2$ et $0\leq z \leq h$. Pour chacun de ces cylindres, on est dans un plan parallèle à $xOy$, c'est à dire qu'on est à $z$ fixé. On connaît l'équation d'un cercle dans un tel plan $x^2+y^2 = r(z)^2$, on en déduit que $x^2 + y^2 = R^2 = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi R^2 \mathrm{d} z $, il ne reste plus qu'à sommer :
\begin{align*}
    V_{\mathrm{cyl}} = \displaystyle \int_{0}^{h} \pi R^2 \, \mathrm{d} z = \pi R^2 h
\end{align*}
\item \textbf{(2 pts)} On décompose la toupie en petit cylindre de hauteur $\mathrm{d} z$ et on  additionne les volumes de tous ces cylindres pour trouver le volume de la toupie. Le volume de chacun des cylindres est $\pi r(z)^2 \mathrm{d} y$.\newline
A $z$ fixé, on connaît l'équation d'un cercle dans le plan $xOy$: $x^2+y^2 = \frac{1}{\sqrt{\abs{z}}}$, on en déduit que $x^2 + y^2 =  \frac{1}{\sqrt{\abs{z}}} = r(z)^2$. Donc, le volume du cylindre pour un $z$ donné est $\pi  \frac{1}{\sqrt{\abs{z}}} \mathrm{d} z $, il ne reste plus qu'à sommer :
\begin{equation}
    V_{\mathrm{toupie}} = \displaystyle \int_{-5}^{5} \pi  \frac{1}{\sqrt{\abs{z}}} \, \mathrm{d} z = 2\displaystyle \int_{0}^{5} \pi  \frac{1}{\sqrt{z}} \, \mathrm{d} z
\end{equation}
Qui est une intégrale généralisée. On trouve finalement
\begin{equation}
    V_{\mathrm{toupie}} = 2\pi\lim \limits_{\epsilon \rightarrow 0}\left[2\sqrt{z} \right]_\epsilon^5 = 4\pi\sqrt{5}
\end{equation}
\end{enumerate}
\section{Exercice 3 \textbf{(4 pts)}}

{1. \textbf{(1 pts)} C'est une équation différentielle linéaire (en
y) ordinaire du premier ordre (seule la première dérivée de y intervient)}\\
{}\\

{2. \textbf{(3 pts)} Méthode de la variation de la constante.}

{On commence par résoudre l'EDO sans second membre
:
\[
\frac{dy_{0}}{dx}+\textrm{sin}\left(x\right)y_{0}=0
\]
c'est-à-dire
\[
\frac{y_{0}'}{y_{0}}=-\textrm{sin}\left(x\right)
\]
ou
\[
\frac{d}{dx}\left(\textrm{ln}\left(y_{0}\right)\right)=\frac{d}{dx}\left(\textrm{cos}\left(x\right)+cte\right)
\]
dont la solution générale s'écrit
\[
y_{0}\left(x\right)=\textrm{e}^{\textrm{cos}\left(x\right)+cte}=A\textrm{e}^{\textrm{cos}\left(x\right)}
\]
avec $A$ une constante réelle que l'on suppose être une fonction
$f\left(x\right)$ dans un second temps. On remplace donc $y\left(x\right)$
par $f\left(x\right)\textrm{e}^{\textrm{cos}\left(x\right)}$ dans
l'EDO et l'on obtient
\[
f'\left(x\right)=\frac{1}{\textrm{cos}^{2}\left(x\right)}
\]
en divisant par $\textrm{e}^{\textrm{cos}\left(x\right)}$ de chaque
côté.}

{Cette équation s'intègre alors facilement en reconnaissant
la dérivée de la fonction tangente à droite :
\[
f\left(x\right)=\textrm{tan}\left(x\right)+C
\]
avec $C$ une constante d'intégration réelle.}

{La solution générale recherchée est donc égale à
\[
y_{sol}\left(x\right)=\left(\textrm{tan}\left(x\right)+C\right)\textrm{e}^{\textrm{cos}\left(x\right)}
\]
et son support de définition est celui de la fonction tangente, c'est-à-dire
l'ensemble des réels privé des nombres $\pi/2$ modulo $\pi$ (relatif).}

\section{Exercice 4 \textbf{(6 pts)}}

{1.\textbf{(1 pts)} C'est une équation différentielle non-linéaire
(en y) du premier ordre.}\\
{}\\

{2.\textbf{(1 pts)} La dérivée de la fonction $\textrm{e}^{y\left(1-y\right)}$
s'écrit :
\[
\left(\textrm{e}^{y\left(1-y\right)}\right)'=y'\left(1-2y\right)\textrm{e}^{y\left(1-y\right)}
\]
}\\
{}\\

{3.\textbf{(3 pts)} Cette dérivée conduit à réexprimer l'EDO comme
\[
\frac{d}{dx}\left(\textrm{e}^{y\left(1-y\right)}\right)=2x
\]
d'où
\[
y\left(1-y\right)=\textrm{ln}\left(x^{2}+cte\right)
\]
La condition $y\left(1\right)=1$ élimine la constante : $cte=0$
puisque $\textrm{ln}\left(1+cte\right)=0$.}

{Le support de définition de $y\left(x\right)$ est
donc au maximum la droite réelle privée du point $0$.}

{Il reste à exprimer $y\left(x\right)$ de façon explicite.
L'équation $y\left(1-y\right)=\textrm{ln}\left(x^{2}\right)$ s'écrit
aussi
\[
y^{2}-y+\textrm{ln}\left(x^{2}\right)=0
\]
qui donne deux solutions a priori possibles
\[
\frac{1\pm\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
dont en fait seule celle avec le signe $+$ correspond à la solution
recherchée car devant vérifier la condition $y\left(1\right)=1$ :
\[
y_{sol}\left(x\right)=\frac{1+\sqrt{1-4\textrm{ln}\left(x^{2}\right)}}{2}
\]
On remarque que le support de définition de $y_{sol}\left(x\right)$
est plus restreint que prévu ci-dessus : il doit correspondre à la
condition de réalité de la racine carrée : $1-4\textrm{ln}\left(x^{2}\right)\geq0$,
c'est-à-dire 
\[
\left|x\right|\leq\textrm{e}^{1/8}
\]
Finalement, l'ensemble de définition de $y_{sol}\left(x\right)$ est
\[
\left[-\textrm{e}^{1/8},0\right[U\left]0,\textrm{e}^{1/8}\right]
\]
}
\section{Exercice 5 \textit{\normalsize{Une ballade en voiture en guise de dessert}{[}5 points{]}}}
\subsubsection*{- \`a l'arr\^et }

\begin{enumerate}
\item \textbf{(0.5 pts)}Cette \'equation diff\'erentielle est lin\'eaire, \`a coefficients constants et elle est du second ordre.
\item \textbf{(1.5 pts)}On \'ecrit le polyn\^ome associ\'e et on cherche ses solutions.
\begin{align*}
mp^{2}+fp+k=0
\end{align*}
avec $\Delta= f^2-4mk$.

Si $\Delta >0$  (c'est \`a dire si $f^{2} > 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent :
\begin{align*}
p_{1,2}=\frac{- f \pm \sqrt{\Delta}}{2m}
\end{align*}
et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= Ae^{p_{1}t}+Be^{p_{2}t}
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.

Si $\Delta =0$ (c'est \`a dire si $f^{2} = 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
\begin{align*}
p_{1} = p_{2} = \frac{-f }{2m}
\end{align*}
et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= (At+B)e^{p_{1}t}
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.

Si $\Delta <0$ (c'est \`a dire si $f^{2} < 4mk$), les solutions du polyn\^ome associ\'e s'\'ecrivent
\begin{align*}
p_{1,2} = \alpha \pm i \omega   
\end{align*}
avec $\alpha= \frac{- f}{2m}$ et $\omega =\frac{\sqrt{-\Delta}}{2m}$

et les solutions de l'\'equation diff\'erentielle s'expriment :
\begin{align*}
y= e^{\alpha t} [A\  \text{sin}(\omega t) + B\  \text{cos} (\omega t)]
\end{align*}
o\`u $A$ et $B$ sont des constantes r\'eelles.
\end{enumerate}

\subsection{- Sur une route "ondul\'ee"}

L'\'equation diff\'erentielle a maintenant un second membre. Elle s'\'ecrit 
\begin{align*}
m\frac{d^{2}y}{dt^{2}}+f \frac{dy}{dt} +ky = F\text{cos} \sqrt{17}t
\end{align*}

\begin{enumerate}
\item \textbf{(0.5 pts)}En effectuant les applications num\'eriques, on obtient :
\begin{align*}
\frac{d^{2}y}{dt^{2}}+2 \frac{dy}{dt} +17y = \text{cos} \sqrt{17}t
\end{align*}
\item \textbf{(1.5 pts)}L'intervalle de d\'efinition de $t$ est ici l'ensemble des r\'eels positifs, la variable $t$ \'etant ici le temps.  Pour obtenir la solution g\'en\'erale de cette \'equation diff\'erentielle, on cherche d'abord la solution g\'en\'erale de l'\'equation sans second membre. On se reporte aux r\'esultats pr\'ec\'edents. Ici $\Delta$ est n\'egatif et on a $\sqrt{-\Delta} = 8$, ce qui conduit \`a $\alpha =-1$ et $\omega=4$. La solution g\'en\'erale de l'\'equation sans second membre s'\'ecrit donc :
\begin{center}

{$y= e^{- t}[A\  \text{sin}(4 t) + B\  \text{cos} (4 t)]$ , A et B \'etant deux r\'eels.}
\end{center}

{On cherche maintenant une solution particuli\`ere de l'\'equation avec second membre, de la forme:}
\begin{center}

{$y= C\ \text{sin}(\sqrt{17} t) + D\  \text{cos} (\sqrt{17} t))$ , C et D \'etant deux r\'eels.}
\end{center}

{Les d\'eriv\'ees $\frac{dy}{dt}$ et $\frac{d^{2}y}{dt^{2}}$ s'\'ecrivent alors:}
\begin{center}

{ $ \frac{dy}{dt}\ = \sqrt{17} C\ \text{cos}(\sqrt{17} t) - \sqrt{17\ } D\ \text{sin} (\sqrt{17} t))$}

{$\frac{d^{2}y}{dt^{2}}= - \ 17 C\ \text{sin}(\sqrt{17} t) \ \  -17\ D\ \text{cos} (\sqrt{17} t))$}
\end{center}

{En rempla\c cant $\frac{d^{2}y}{dt^{2}}$, $\frac{dy}{dt}$ et $y$ dans l'\'equation diff\'erentielle compl\`ete, on obtient en r\'eunissant les termes en sinus d'un c\^ot\'e et ceux en cosinus de l'autre:}

\begin{center}
{$[- 17C -2 \sqrt{17} D+17C]\ \text{sin}(\sqrt{17} t) + [-17 D+2\sqrt{17} C +17D]\ \text{cos}(\sqrt{17} t) = \text{cos}( \sqrt{17} t)$}
\end{center}

{Par identification des termes en $\text{sin}(\sqrt{17} t)$ et ceux en $\text{cos}(\sqrt{17} t)$ de part et d'autre de l'\'egalit\'e, on obtient :}

\begin{center}
{$D=0$ et $C=\frac{1}{2\sqrt{17}}$.}
\end{center}

{Soit une solution particuli\`ere de l'\'equation avec second membre \'egale \`a:}

\begin{center}
{$y= \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$,}
\end{center}

{et une solution g\'en\'erale de l'\'equation compl\`ete avec second membre :}

\begin{center}
{$y= e^{- t} [A\  \text{sin}(4 t) + B\  \text{cos}(4 t)] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$, A et B \'etant deux r\'eels}
\end{center}
\item \textbf{(1 pts)} Pour trouver la solution particuli\`ere de l'\'equation compl\`ete avec second membre, associ\'ee aux conditions aux limites $y(t=0)=1$ et $y'(t=0)=-1$ on calcule $y(t=0)$ et $y'(t=0)$:

\begin{center}
{$y(t=0)= B+ \frac{1}{2\sqrt{17}}*0=1$, }

{$y'(t=0)= [4A - B + \frac{1}{2}]=-1$.}
\end{center}

{On obtient donc $B=1$ et $A=-\frac{1}{8}$ et la solution s''ecrit :}

\begin{center}
{$y= e^{- t} [-\frac{1}{8} \  \text{sin}(4 t) +  \text{cos}(4 t) ] + \frac{1}{2\sqrt{17}}\ \text{sin}(\sqrt{17} t)$.}
\end{center}
\end{enumerate}

\end{document}

>>>>

\documentclass[fleqn]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{mathexam}
\usepackage[french]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

@@

\begin{document}
\ExamInstrBox{
\begin{center}
\textbf{Polycopié de cours autorisé. \\ Calculatrice et TD interdits}
\end{center}
}

\section{Exercice 1 : intégrales (45')}
\subsection{Calcul d'intégrales (3x10')}
Calculer les intégrales suivantes et préciser en une phrase si elles sont généralisées ou non :
\begin{enumerate}
   \item ${\displaystyle \int_0^1 x^2 \mathrm{ln}(x) dx}$ 
   \item ${\displaystyle \int_0^{\pi/2} \dfrac{\cos(x)}{\sqrt{1+3\sin(x)}}dx}$
   \item ${\displaystyle \int_{-\pi/2}^{\pi/2} \dfrac{\cos(x)-1}{\sin^2(x)}dx}$
\end{enumerate}
\end{document}
